---
title: "rivr (compiled)"
author: ""
date: "June 30, 2017"
output: html_document
---

<style>
  .col2 {
    columns: 2 200px;         /* number of columns and width in pixels*/
    -webkit-columns: 2 200px; /* chrome, safari */
    -moz-columns: 2 200px;    /* firefox */
  }
</style>

```{r apriori, include=FALSE}
path <- "/home/kondziu/workspace/R-dyntrace/data/rivr.sqlite"
```

```{r setup, include=FALSE}
source("/home/kondziu/workspace/R-dyntrace/report-basic.r")
```

Functions: `r n.functions`  
Calls: `r n.calls`  
Promises: `r n.promises` (+ `r n.alien.promises` alien)  
Promise forces: `r n.promise.forces` (+ `r n.alien.promise.forces` alien)  
Promise lookups: `r n.promise.lookups` (+ `r n.alien.promise.lookups` alien)  

## Promises

Alien promises are promises which were created before we began tracing. We don't have all the information regarding them, so we generally ignore them, including in the data below. They also are a subject of a lot of lookups in functions from base::*, which we don't record to save disk space.

### Promise evaluations
<div class="col2">
``` {r evaluations-print, echo=FALSE, , comment=""}
force.histogram <- get_force_histogram()
#lookup.histogram <- get_lookup_histogram()
promise.evaluation.histogram <- get_promise_evaluation_histogram()
#force.histogram
#lookup.histogram
#promise.evaluation.histogram

slices <- force.histogram$number
labels <- force.histogram$no.of.forces
par(mar=c(0,2,2,2))
pie(slices, labels, main="No. of forces")

#labels <- lookup.histogram$no.of.lookups
#values <- lookup.histogram$number
#par(mar=c(2,2,2,2))
#barplot(values, names.arg=labels, space=0, main="No. of lookups")

labels <- promise.evaluation.histogram$no.of.evluations
values <- promise.evaluation.histogram$number
par(mar=c(2,2,2,2))
barplot(values, names.arg=labels, space=0, main="No. of evaluations")
```
</div>

### Promise lifestyles



<div class="col2">
``` {r lifestyles-print, echo=FALSE, , comment="", warning=FALSE}
promise.lifestyles <- get_lifestyles()
promise.lifestyles
```

**immediate local** - evaluated in same call as it originated (effective distance = 0 and actual distance = 0)  
**local** - same, but passed to primitive or specials in origin (effective distance = 0, actual distance > 0)  
**immediate branch-local** - evaluated in a child call to its origin (effective distance = 1)  
**branch-local** - same, but further descendant  
**escaped** - evaluated in a call to which there is no path from its origin  
**virgin** - promise was never evaluated

``` {r lifestyles-pie, echo=FALSE, fig.width=6, fig.height=5.5}
slices <- (promise.lifestyles %>% arrange(number))$number
labels <- (promise.lifestyles %>% arrange(number) %>% mutate(lifestyle = ifelse(percent>1, lifestyle, "")))$lifestyle
par(mar=c(0,2,2,2))
pie(slices, labels, main="Promise lifestyles")
```
</div>

### Promise evaluation distances

Effective distance is the distance from the call where a promise originated to the call in which it was evaluated. It is measured along the call graph (along inbound edges).  Since the data is gathered from a trace, the call graph is a tree.

Special values:  
**NA** - promise was not executed, there is no distance measure (a virgin promise)  
**-1** - promise was executed but there is no path from the origin to the point of evaluation (an escaped promise)    
``` {r effective-distances, echo=FALSE, fig.width=12, fig.height=6}
histogram <- get_effective_distances()
labels <- histogram$effective_distance
values <- histogram$number
par(mar=c(2,2,2,2))
barplot(values, names.arg=labels, space=0, main="Effective distances")
```

``` {r actual-distances, echo=FALSE, fig.width=12, fig.height=6}
histogram <- get_actual_distances()
labels <- histogram$effective_distance
values <- histogram$number
par(mar=c(2,2,2,2))
barplot(values, names.arg=labels, space=0, main="Actual distances")
```

### Promise code types

Promise types follow the type designations in``Rinternals.h``.
**α→β** indicates that type **α** variously evaluates to type **β**.
Specifically **PROM→α** indicates that it's a promise containing the expression of type **α**. 
**BCODE →α** indicates that it's a byte-compiled expression that was originally of type **α**.
**SYM→α** indicates that it's a symbol that points to an expression of type **α**.

<div class="col2">
``` {r promise-types-print, echo=FALSE, comment=""}
promise.types <- get_promise_types()
promise.full.types <- get_full_promise_types()
promise.types
promise.full.types
```

``` {r promise-types-pie, echo=FALSE, fig.width=5, fig.height=4.5}
slices <- (promise.types %>% arrange(number))$number
labels <- (promise.types %>% arrange(number) %>% mutate(type = ifelse(percent>1, type, "")))$type
par(mar=c(0,2,2,2))
pie(slices, labels, main="Basic types")

slices <- (promise.full.types %>% arrange(number))$number
labels <- (promise.full.types %>% arrange(number) %>% mutate(type = ifelse(percent>1, type, "")))$type
par(mar=c(0,2,2,2))
pie(slices, labels, main="Full types")
```
</div>

## Functions and Calls

### Strictness



### Calls to specific functions

```{r function-calls-selected-counts, echo=FALSE, comment=""}
  get_function_calls("^<anonymous>$", "eval", "force", "delayAssign", "return") %>% data.frame
```




