<style>
  .col2 {
    columns: 2 200px;         /* number of columns and width in pixels*/
    -webkit-columns: 2 200px; /* chrome, safari */
    -moz-columns: 2 200px;    /* firefox */
  }
  
   div.together {
    display: inline-block;
    width: 100%;
 }
</style>

```{r setup, include=FALSE}
library(ggplot2)
path <- "%%PATH%%"
print_data <- TRUE
source("%%FUNCTIONS%%")
```

```{r evaluation-order-heavy-lifting, include=FALSE}
call.evaluation.order <- get_call_promise_evaluation_order()
function.evaluation.order <- get_function_promise_evaluation_order(call.evaluation.order)
promise.force.order.histogram <- get_function_promise_force_order_histogram(function.evaluation.order, cutoff=10)
```

<div class="col2">
## Basic Information
Dataset: `r path`  
Functions: `r pp(n.functions)`  
Calls: `r pp(n.calls)`  
Promises: `r pp(n.promises)` (+ `r pp(n.alien.promises)` alien)  
Promise forces: `r pp(n.promise.forces)` (+ `r pp(n.alien.promise.forces)` alien)  
Promise lookups: `r pp(n.promise.lookups)` (+ `r pp(n.alien.promise.lookups)` alien)  

<div class="together">
## Metadata
```{r metadata, echo=FALSE, comment=""}
print(metadata %>% as.data.frame %>% unique, row.names=FALSE, use.names=FALSE)
```
</div>

</div>

## Promises

Alien promises are promises which were created before we began tracing. We don't have all the information regarding them, so we generally ignore them, including in the data below. They also are a subject of a lot of lookups in functions `from base::*`, which we don't record to save disk space.

### Promise evaluations

When a promise is evaluated it is either **forced**, if it was not evaluated before, or **looked up**, if it was previously evaluated.

<div class="col2">
<div class="together">
``` {r evaluations-forces, echo=FALSE, , comment="", fig.width=7, fig.height=5}
force.histogram <- get_force_histogram()

slices <- force.histogram$number
labels <- force.histogram$no.of.forces
par(mar=c(0,2,2,2))

pie(slices, labels, main="No. of forces")
if(print_data) print(force.histogram %>% as.data.frame %>% transform(number=pp(number), percent=pp(percent)), row.names=FALSE)
```
</div>

``` {r evaluations-forces-cat, echo=FALSE, , comment="", fig.width=7, fig.height=5}
fuzzy.force.histogram <- get_fuzzy_force_histogram()

slices <- rev(fuzzy.force.histogram$number)
labels <- rev(fuzzy.force.histogram$classification)
par(mar=c(0,2,2,2))

pie(slices, labels, main="Handling of promises")
if(print_data) print(fuzzy.force.histogram %>% as.data.frame %>% transform(number=pp(number), percent=pp(percent)), row.names=FALSE)
```

</div>

<center>
<!--div class="together"-->
``` {r evaluations-total, echo=FALSE, , comment="", fig.width=12, fig.height=5}
cutoff <- 10
promise.evaluation.histogram <- get_promise_evaluation_histogram(cutoff=cutoff)

promise.evaluation.histogram_with_labels <-
  promise.evaluation.histogram %>%
  mutate(no.of.evaluations=
           ifelse(no.of.evaluations==Inf,
                  paste(">",cutoff, sep=""),
                  as.character(no.of.evaluations)))
labels <- promise.evaluation.histogram_with_labels$no.of.evaluations
values <- promise.evaluation.histogram$number
par(mar=c(2,2,2,2))

barplot(values, names.arg=labels, space=0, main="No. of evaluations (forces + lookups)")
if(print_data) print(promise.evaluation.histogram_with_labels %>% as.data.frame %>% transform(number=pp(number), percent=pp(percent)), row.names=FALSE)
```
<!--/div-->
</center>

### Promise code types

Promise types follow the type designations in``Rinternals.h``.
**α→β** indicates that type **α** variously evaluates to type **β**.
Specifically **PROM→α** indicates that it's a promise containing the expression of type **α**. 
**BCODE →α** indicates that it's a byte-compiled expression that was originally of type **α**.
**SYM→α** indicates that it's a symbol that points to an expression of type **α**.

<div class="col2">
<div class="together">
``` {r promise-types-pie, echo=FALSE, comment="", fig.width=5, fig.height=4.5}
promise.types <- get_promise_types(cutoff=NA)

slices <- (promise.types %>% arrange(number))$number
labels <- (promise.types %>% arrange(number) %>% mutate(type = ifelse(percent>1, type, "")))$type
par(mar=c(0,2,2,2))

pie(slices, labels, main="Basic types")
#barplot(slices, names.arg=labels, space=0, main="Basic types")
if(print_data) print(promise.types %>% as.data.frame %>% transform(number=pp(number), percent=pp(percent)), row.names=FALSE)
```
</div>

<!--div class="together"-->
``` {r promise-full-types-pie, echo=FALSE, fig.width=5, fig.height=4.5, comment=""}
promise.full.types <- get_full_promise_types(cutoff=1)

slices <- (promise.full.types %>% arrange(number))$number
labels <- (promise.full.types %>% arrange(number) %>% mutate(type = ifelse(percent>1, type, "")))$type
par(mar=c(0,2,2,2))

pie(slices, labels, main="Full types")
#barplot(slices, names.arg=labels, space=0, main="Full types") + theme(axis.text.x = element_text(angle = 90, hjust = 1))
if(print_data) print(promise.full.types %>% as.data.frame %>% transform(number=pp(number), percent=pp(percent)), row.names=FALSE)
```
<!--/div-->
</div>


### Promise return types

<div class="col2">
<div class="together">
``` {r promise-return-types-pie, echo=FALSE, comment="", fig.width=5, fig.height=4.5}
return.types <- get_promise_return_types(cutoff=NA)

slices <- (return.types %>% arrange(number))$number
labels <- (return.types %>% arrange(number) %>% mutate(type = ifelse(percent>1, type, "")))$type
par(mar=c(0,2,2,2))

pie(slices, labels, main="Return types")
#barplot(slices, names.arg=labels, space=0, main="Basic types")
if(print_data) print(return.types %>% as.data.frame %>% transform(number=pp(number), percent=pp(percent)), row.names=FALSE)
```
</div>

``` {r promise-bcode-return-types-pie, echo=FALSE, comment="", fig.width=5, fig.height=4.5}
return.types <- get_promise_return_types_by_type(cutoff=NA, promise_type="BCODE")

slices <- (return.types %>% arrange(number))$number
labels <- (return.types %>% arrange(number) %>% mutate(type = ifelse(percent>1, type, "")))$type
par(mar=c(0,2,2,2))

pie(slices, labels, main="Return types for BCODE")
#barplot(slices, names.arg=labels, space=0, main="Basic types")
if(print_data) print(return.types %>% as.data.frame %>% transform(number=pp(number), percent=pp(percent)), row.names=FALSE)
```

</div>


### Promise forces by types

<center>
``` {r evaluations-forces-by-type, echo=FALSE, , comment="", fig.width=7, fig.height=10}
force.histogram.by.type <- get_force_histogram_by_type()

par(mar=c(1,1,1,1))

ggplot(
  force.histogram.by.type, 
  aes(x=type, y=number, fill=factor(no.of.forces))) + 
geom_bar(stat="identity", position="dodge") +
scale_fill_discrete(name="No. of forces",
                    breaks=(force.histogram.by.type$no.of.forces %>% unique),
                    labels=(force.histogram.by.type$no.of.forces %>% unique)) +
xlab("") +
ylab("") + 
ggtitle("No. of forces by type") +
theme(plot.title = element_text(face="bold", hjust = 0.5))

if(print_data) 
  print(force.histogram.by.type %>% 
          as.data.frame %>% 
          transform(
            number=pp(number), 
            percent_within_type=pp(percent_within_type), 
            percent_overall=pp(percent_overall)), 
        row.names=FALSE)
```
</center>

### Promise Lifestyles


``` {r lifestyles-print, echo=FALSE, , comment="", warning=FALSE, fig.width=8, fig.height=5}

```

<!--div class="col2"-->

**immediate local** - evaluated in same call as it originated (effective distance = 0 and actual distance = 0)  
**local** - same, but passed to primitive or specials in origin (effective distance = 0, actual distance > 0)  
**immediate branch-local** - evaluated in a child call to its origin (effective distance = 1)  
**branch-local** - same, but further descendant  
**escaped** - evaluated in a call to which there is no path from its origin  
**virgin** - promise was never evaluated

<center>
``` {r lifestyles-pie, echo=FALSE, fig.width=10, fig.height=5.5, comment="", warning=FALSE}
promise.lifestyles <- get_lifestyles()

slices <- (promise.lifestyles %>% arrange(number))$number
labels <- (promise.lifestyles %>% arrange(number) %>% mutate(lifestyle = ifelse(percent>1, lifestyle, "")))$lifestyle
par(mar=c(0,5,2,5))

pie(slices, labels, main="Promise lifestyles")
if (print_data) 
  print(
    promise.lifestyles %>% 
      as.data.frame %>% 
      transform(
        number=pp(number),
        percent=pp(percent)), 
    row.names=FALSE)
```
</center>
<!--/div-->

### Promise Evaluation Distances

Effective distance is the distance from the call where a promise originated to the call in which it was evaluated. It is measured along the call graph (along inbound edges).  Since the data is gathered from a trace, the call graph is a tree.

Special values:  
**NA** - promise was not executed, there is no distance measure (a virgin promise)  
**-1** - promise was executed but there is no path from the origin to the point of evaluation (an escaped promise)    

<div class="col2">
<div class="together">
``` {r effective-distances, echo=FALSE, fig.width=7, fig.height=5, comment=""}
cutoff <- 10
histogram <- get_effective_distances(cutoff=cutoff)
histogram_with_labels <- 
  histogram %>% 
  mutate(effective_distance=
           ifelse(effective_distance==Inf, 
                  paste(">",cutoff, sep=""),
                  as.character(effective_distance)))
labels <- histogram_with_labels$effective_distance
values <- histogram$number
par(mar=c(2,2,2,2))

barplot(values, names.arg=labels, space=0, main="Effective distances")
if(print_data) print(histogram_with_labels %>% as.data.frame %>% transform(number=pp(number), percent=pp(percent)), row.names=FALSE)

# TODO How many promises change types between creation and evaluation????
```
</div>

<!--div class="together"-->
``` {r actual-distances, echo=FALSE, fig.width=7, fig.height=5, comment=""}
cutoff <- 10
histogram <- get_actual_distances(cutoff=cutoff)
histogram_with_labels <- 
  histogram %>% 
  mutate(actual_distance=
           ifelse(actual_distance==Inf, 
                  paste(">", cutoff, sep=""), 
                  as.character(actual_distance)))
labels <- histogram_with_labels$actual_distance
values <- histogram$number
par(mar=c(2,2,2,2))

barplot(values, names.arg=labels, space=0, main="Actual distances")
if(print_data) print(histogram_with_labels %>% as.data.frame %>% transform(number=pp(number), percent=pp(percent)), row.names=FALSE)
```
<!--/div-->
</div>

## Functions and Calls

### Functions and Calls by Type

<div class="col2">
<div class="together">
```{r calls-by-type, echo=FALSE, fig.width=5, fig.height=4.5, comment=""}
calls.types <- get_calls_by_type()

slices <- (calls.types %>% arrange(number))$number
labels <- (calls.types %>% arrange(number) %>% mutate(type = ifelse(percent>1, type, "")))$type
par(mar=c(0,2,2,2))

pie(slices, labels, main="Call types")
if(print_data) print(calls.types %>% as.data.frame %>% transform(number=pp(number), percent=pp(percent)), row.names=FALSE)
```
</div>

```{r functions-by-type, echo=FALSE, fig.width=5, fig.height=4.5, comment=""}
calls.functions <- get_functions_by_type()

slices <- (calls.functions %>% arrange(number))$number
labels <- (calls.functions %>% arrange(number) %>% mutate(type = ifelse(percent>1, type, "")))$type
par(mar=c(0,2,2,2))

pie(slices, labels, main="Function types")
if(print_data) print(calls.functions %>% as.data.frame %>% transform(number=pp(number), percent=pp(percent)), row.names=FALSE)
```
</div>

### Byte-code Compilation

When calling a function it's either byte-code **compiled** or **uncompiled**. When agregated across all calls to a function, a function can be:
**compiled** - all calls to this function are compiled,
**uncompiled** - none of the calls to this function are compiled,
**after 1st** - a function can be compiled by the JIT, so it can be uncompiled initially and then compiled later on,
**erratic** - a function is sometimes compiled and sometimes not; it can indicate that two separate functions are indistinguishable to the tracer (same definition, unknown location) and one of them is compiled while another is not.


<div class="col2">
``` {r compiled-calls-closures, echo=FALSE, comment="", fig.width=5, fig.height=4.5}
closure.call.compilation.histogram <- get_call_compilation_histogram_by_type("closure") %>% rename(percent = percent_within_type)

slices <- (closure.call.compilation.histogram)$number
labels <- (closure.call.compilation.histogram)$compiled
par(mar=c(0,2,2,2))

pie(slices, labels, main="Compiled closures (calls)")
if(print_data) print(closure.call.compilation.histogram %>% as.data.frame %>% transform(number=pp(number), percent=pp(percent)), row.names=FALSE)
```
<div class="together">

</div>
<!--div class="together"-->
``` {r compiled-functions-closures, echo=FALSE, comment="", fig.width=5, fig.height=4.5}
closure.function.compilation.histogram <- get_function_compilation_histogram_by_type_actual("closure") %>% rename(percent = percent_within_type)

slices <- (closure.function.compilation.histogram)$number
labels <- (closure.function.compilation.histogram %>% mutate(type = ifelse(percent>1, compiled, "")))$compiled
par(mar=c(0,2,2,2))

pie(slices, labels, main="Compiled closures (functions)")
if(print_data) print(closure.function.compilation.histogram %>% as.data.frame %>% transform(number=pp(number), percent=pp(percent)), row.names=FALSE)
```
<!--/div-->
</div>

<!--
### Recursion

Determining whether a function is recursive or not from a trace is necessarily imprecise. A call ``c`` to some function ``f`` is recursive if there is at least one more call to ``f`` above ``c`` in the stack trace. Instead of talking about recursive functions, we just show what percentage of their calls is recursive under this definition.

<div class="col2">
``` {r recursion-data, echo=FALSE, include=FALSE}
# recursion.info <- traverse_call_tree()
# recursive.calls.histogram <- get_call_recursion_histogram(recursion.info)
# recursive.functions.histogram <- get_function_recursion_histogram(recursion.info)

```

``` {r recursive-calls, echo=FALSE, comment="", fig.width=5, fig.height=4.5}
# par(mar=c(0,2,2,2))
# pie(recursive.calls.histogram$number, recursive.calls.histogram$recursive, main="Recursive calls")
# if(print_data) print(recursive.calls.histogram %>% as.data.frame %>% transform(number=pp(number), percent=pp(percent)), row.names=FALSE)
```
<div class="together">

</div>
``` {r recursive-functions, echo=FALSE, comment="", fig.width=5, fig.height=4.5}
# par(mar=c(2,2,2,2))
# barplot(recursive.functions.histogram$number, names.arg=recursive.functions.histogram$recursion_rate, space=0, main="Recursive call rates by function")
# if(print_data) print(recursive.functions.histogram %>% as.data.frame %>% transform(number=pp(number), percent=pp(percent)), row.names=FALSE)
```
</div>
-->

### Interaction Among Promises

<!-- **Free** promises are promises that are not forced by another promise. **Non-free** promises are forced by some other promise. -->
<!-- A force **cascade** is a situation where one promise forces another promise. This shows how many promise evaluations caused a cascade. I exclude the topmost promise. -->

Promises forced by other promises: 
`TRUE` is when a promise was created within some promise x but was forced within a different promise y. 
`FALSE` is when a promise was forced within the same promise as it was created. 
`NA` means the promise was not forced.

<div class="col2">
<div class="together">

``` {r promise-forced-by-another-evaluation, echo=FALSE, comment="", fig.width=5, fig.height=4.5}
histogram <- get_promises_forced_by_another_evaluation_histogram()
par(mar=c(0,2,2,2))
pie(histogram$number, (histogram$forced_by_another %>% make_labels), main="Promises forced by another promise")
if(print_data) print(histogram %>% as.data.frame %>% transform(number=pp(number), percent=pp(percent)), row.names=FALSE)
```

</div>

``` {r promise-cascades, echo=FALSE, comment="", fig.width=5, fig.height=4.5}
histogram <- get_cascading_promises_histogram()
labels <- histogram$number_of_forced_promises
values <- histogram$number
par(mar=c(2,2,2,2))

barplot(values, names.arg=labels, space=0, main="Promises forcing other promises")
if(print_data) print(histogram %>% as.data.frame %>% transform(number=pp(number), percent=pp(percent)), row.names=FALSE)
```

</div>

### Promises per Call and per Function

### Strictness

``` {r strictness-data, echo=FALSE, include=FALSE}
strict.calls.histogram <- get_call_strictness()
strict.calls.ratio.histogram <- get_call_strictness_ratios()
strict.calls.rate.histogram <- get_call_strictness_rate()
strict.calls.by.type.histogram <- get_call_strictness_by_type()
strict.functions.histogram <- get_function_strictness()
strict.functions.rate.histogram <- get_function_strictness_rate()
strict.functions.by.type.histogram <- get_function_strictness_by_type()
#strict.arguments.histogram <- 
```

A call is **strict** if all of its arguments are **evaluated** and none of the arguments is **escaped**. (This means that a function may pass unevaluated promises to a child call and still be strict, but it cannot leak the promise).
A function is **strict** if all of its calls are **strict**. Functions that don't take promises are neither **strict** nor **non-strict** -- these can be either arrity-0 closures or built-in (which accept values rather than promises).

**Strictness ratio** of the form `n/m` means that a call evaluated `n` promises out of `m` promises that were created for it. `0/0` can mean either that a function had no arguments or that it was a built-in etc. (had arguments, but they were passed without promises).

<div class="col2">
<div class="together">

``` {r strict-calls, echo=FALSE, comment="", fig.width=5, fig.height=4.5}
par(mar=c(0,2,2,2))
pie(strict.calls.histogram$number, (strict.calls.histogram$strict %>% make_labels), main="Call strictness")
if(print_data) print(strict.calls.histogram %>% as.data.frame %>% transform(number=pp(number), percent=pp(percent)), row.names=FALSE)
```

</div>

``` {r strict-by-types, echo=FALSE, comment="", fig.width=5, fig.height=4.5}
par(mar=c(1,1,1,1))

ggplot(
  strict.calls.by.type.histogram, 
  aes(x=type, y=number, fill=factor(strict))) + 
geom_bar(stat="identity", position="dodge") +
scale_fill_discrete(name="Strictness",
                    breaks=(strict.calls.by.type.histogram$strict %>% unique),
                    labels=(strict.calls.by.type.histogram$strict %>% unique)) +
xlab("") +
ylab("") + 
ggtitle("Call strictness by type") +
theme(plot.title = element_text(face="bold", hjust = 0.5))

if(print_data) 
  print(strict.calls.by.type.histogram %>% 
        as.data.frame %>% 
        transform(
          number=pp(number), 
          percent=pp(percent)), 
        row.names=FALSE)
```

``` {r strict-rate, echo=FALSE, comment="", fig.width=5, fig.height=4.5}
# par(mar=c(2,2,2,2))
# barplot(strict.calls.rate.histogram$number, names.arg=(strict.calls.rate.histogram$strictness_rate %>% make_labels), space=0, main="Call strictness (rate of evaluated promises)")
# if(print_data) print(strict.calls.rate.histogram %>% as.data.frame %>% transform(number=pp(number), percent=pp(percent)), row.names=FALSE)
```

</div>

<center>
``` {r strict-calls-ratio, echo=FALSE, comment="", fig.width=12, fig.height=4.5}
par(mar=c(2,2,2,1))
barplot(strict.calls.ratio.histogram$number, names.arg=(strict.calls.ratio.histogram$strictness_ratio), space=0, main="Call strictness (ratio of evaluated promises)", cex.names=0.5)
if(print_data) print(strict.calls.ratio.histogram %>% as.data.frame %>% transform(number=pp(number), percent=pp(percent)), row.names=FALSE)
````
</center>

<div class="col2">
<div class="together">
``` {r strict-functions, echo=FALSE, comment="", fig.width=5, fig.height=4.5}
par(mar=c(0,2,2,2))
pie(strict.functions.histogram$number, (strict.functions.histogram$strict %>% make_labels), main="Function strictness")
if(print_data) print(strict.functions.histogram %>% as.data.frame %>% transform(number=pp(number), percent=pp(percent)), row.names=FALSE)
```
</div>
``` {r strict-functions-by-types, echo=FALSE, comment="", fig.width=5, fig.height=4.5}
par(mar=c(1,1,1,1))

ggplot(
  strict.functions.by.type.histogram, 
  aes(x=type, y=number, fill=factor(strict))) + 
geom_bar(stat="identity", position="dodge") +
scale_fill_discrete(name="Strictness",
                    breaks=(strict.functions.by.type.histogram$strict %>% unique),
                    labels=(strict.functions.by.type.histogram$strict %>% unique)) +
xlab("") +
ylab("") + 
ggtitle("Function strictness by type") +
theme(plot.title = element_text(face="bold", hjust = 0.5))

if(print_data) 
  print(strict.functions.by.type.histogram %>% 
        as.data.frame %>% 
        transform(
          number=pp(number), 
          percent=pp(percent)), 
        row.names=FALSE)
```
</div>

<center>
``` {r strict-functions-rate, echo=FALSE, comment="", fig.width=12, fig.height=4.5}
par(mar=c(2,2,2,2))
barplot(strict.functions.rate.histogram$number, names.arg=(strict.functions.rate.histogram$strictness_rate %>% make_labels), space=0, main="Function strictness (rate of strict calls)")
if(print_data) print(strict.functions.rate.histogram %>% as.data.frame %>% transform(number=pp(number), percent=pp(percent)), row.names=FALSE)
````
</center>

### Evaluation order

``` {r force-order, echo=FALSE, , comment="", fig.width=12, fig.height=5}
cutoff <- 10

promise.force.order.histogram_with_labels <-
  promise.force.order.histogram %>%
  mutate(no.of.force.orders=
           ifelse(is.na(no.of.force.orders), 
                  0,
                  ifelse(no.of.force.orders==Inf,
                         paste(">",cutoff, sep=""),
                         as.character(no.of.force.orders))))

labels <- promise.force.order.histogram_with_labels$no.of.force.orders
values <- promise.force.order.histogram$number
par(mar=c(2,2,2,2))

barplot(values, names.arg=labels, space=0, main="No. of force orders")
if(print_data) print(promise.force.order.histogram_with_labels %>% as.data.frame %>% transform(number=pp(number), percent=pp(percent)), row.names=FALSE)
```

### Most called functions

### Calls to specific functions

```{r function-calls-selected-counts, echo=FALSE, comment=""}
get_function_calls("^<anonymous>$", "eval", "force", "delayAssign", "return", "<-", "<<-", "print") %>% data.frame %>% transform(number=pp(number), percent=pp(percent)) %>% print(., row.names=FALSE)
```




