---
title: "A Look at Vignettes"
author: "Konrad Siek"
date: "22 August 2017"
output: html_document
---

<style>
  .col2 {
    columns: 2 200px;         /* number of columns and width in pixels*/
    -webkit-columns: 2 200px; /* chrome, safari */
    -moz-columns: 2 200px;    /* firefox */
  }
  
   div.together {
    display: inline-block;
    width: 100%;
 }
</style>

```{r setup, include=FALSE}
library(ggplot2)
library(tibble)
library(dplyr)

pp <- function(number) format(number, big.mark=",", scientific=FALSE, trim=FALSE, digits=2)

#summary.vignettes <- readRDS("summary_vignettes.rds")
#summary.packages <- readRDS("summary_packages.rds")

summary.vignettes <- readRDS("~/workspace/R-dyntrace/summary_vignettes.rds")
summary.packages <- readRDS("~/workspace/R-dyntrace/summary_packages.rds")

top.100.packages <- scan("~/workspace/R-dyntrace/top_100_CRAN_packages.txt", as.character())
```

## Basic information

Number of packages: `r length(summary.packages$package)`  
Number of vignettes: `r length((summary.vignettes %>% filter(!is.na(index)))$name)`

## Statistics

<div class="col2">
<div class="together">
```{r has-vignettes, echo=FALSE, comment="", fig.width=6, fig.height=6}
histogram <- summary.packages %>% mutate(has.vignettes=n.vignettes>0) %>% group_by(has.vignettes) %>% summarise(n.packages=n()) %>% mutate(percent=100*n.packages/sum(n.packages)) %>% arrange(n.packages)

labels <- histogram$has.vignettes
values <- histogram$n.packages
#par(mar=c(2,2,2,2))

pie(values, labels, main="Package has vignettes")
print(histogram %>% as.data.frame %>% transform(n.packages=pp(n.packages), percent=pp(percent)), row.names=FALSE)
```
</div>

```{r has-runnable-code, echo=FALSE, comment="", fig.width=6, fig.height=6}
histogram <- 
  summary.packages %>% 
  mutate(runnable.code=ifelse(n.vignettes == 0, "no vignettes",
                              ifelse(n.lines == 0, "no code",
                                     ifelse(n.lines <= 10, "≤ 10 lines",
                                            ifelse(n.lines <= 100, "≤ 100 lines", "> 100 lines"))))) %>%
  group_by(runnable.code) %>% summarise(n.packages=n()) %>% mutate(percent=100*n.packages/sum(n.packages)) %>% 
  arrange(n.packages)

labels <- histogram$runnable.code
values <- histogram$n.packages
#par(mar=c(2,2,2,2))

pie(values, labels, main="Package has runnable code in vignettes")
print(histogram %>% as.data.frame %>% transform(n.packages=pp(n.packages), percent=pp(percent)), row.names=FALSE)
```

</div>

```{r n-vignettes, echo=FALSE, comment="", fig.width=12, fig.height=5}
histogram <- summary.packages %>% group_by(n.vignettes) %>% summarise(n.packages=n()) %>% mutate(percent=100*n.packages/sum(n.packages))

labels <- histogram$n.vignettes
values <- histogram$n.packages
#par(mar=c(2,2,2,2))

barplot(values, names.arg=labels, space=0, main="Number of vignettes per package")
print(histogram %>% as.data.frame %>% transform(n.packages=pp(n.packages), percent=pp(percent)), row.names=FALSE)
```

## Top 100 packages

```{r top-100, echo=FALSE, comment="", fig.width=12, fig.height=5}
summary.top.100.packages <- tibble(package=top.100.packages) %>% left_join(summary.packages, by="package")
summary.top.100.vignettes <- tibble(package=top.100.packages) %>% left_join(summary.vignettes, by="package")
```

### Basic information

Number of packages: `r length(summary.top.100.packages$package)`  
Number of these actually installed: `r length((summary.top.100.packages %>% filter(!is.na(n.vignettes)))$package)`  
Number of vignettes: `r length((summary.top.100.vignettes %>% filter(!is.na(index)))$name)`  

### Statistics

<div class="col2">
<div class="together">
```{r top-100-has-vignettes, echo=FALSE, comment="", fig.width=6, fig.height=6}
histogram <- summary.top.100.packages %>% mutate(has.vignettes=n.vignettes>0) %>% group_by(has.vignettes) %>% summarise(n.packages=n()) %>% mutate(percent=100*n.packages/sum(n.packages)) %>% arrange(n.packages)

labels <- as.logical(histogram$has.vignettes)
values <- histogram$n.packages
#par(mar=c(2,2,2,2))

pie(values, labels, main="Package has vignettes")
print(histogram %>% as.data.frame %>% transform(n.packages=pp(n.packages), percent=pp(percent)), row.names=FALSE)
```
</div>

```{r top-100-has-runnable-code, echo=FALSE, comment="", fig.width=6, fig.height=6}
histogram <- 
  summary.top.100.packages %>% 
  mutate(runnable.code=ifelse(n.vignettes == 0, "no vignettes",
                              ifelse(n.lines == 0, "no code",
                                     ifelse(n.lines <= 10, "≤ 10 lines",
                                            ifelse(n.lines <= 100, "≤ 100 lines", "> 100 lines"))))) %>%
  group_by(runnable.code) %>% summarise(n.packages=n()) %>% mutate(percent=100*n.packages/sum(n.packages)) %>% 
  arrange(n.packages)

labels <- histogram$runnable.code
values <- histogram$n.packages
#par(mar=c(2,2,2,2))

pie(values, labels, main="Package has runnable code in vignettes")
print(histogram %>% as.data.frame %>% transform(n.packages=pp(n.packages), percent=pp(percent)), row.names=FALSE)
```

</div>

```{r top-100-n-vignettes, echo=FALSE, comment="", fig.width=12, fig.height=5}
histogram <- summary.top.100.packages %>% group_by(n.vignettes) %>% summarise(n.packages=n()) %>% mutate(percent=100*n.packages/sum(n.packages))

labels <- histogram$n.vignettes
values <- histogram$n.packages
#par(mar=c(2,2,2,2))

barplot(values, names.arg=labels, space=0, main="Number of vignettes per package")
print(histogram %>% as.data.frame %>% transform(n.packages=pp(n.packages), percent=pp(percent)), row.names=FALSE)
```

### Data (top 100)

```{r top-100-data-packages, echo=FALSE, comment=""}
print(
   summary.top.100.packages %>% 
    as.data.frame %>% 
    transform(n.vignettes=pp(n.vignettes), n.lines=pp(n.lines)), 
  row.names=FALSE)
```

```{r top-100-data-vignettes, echo=FALSE, comment=""}
print(
   summary.top.100.vignettes %>% 
    as.data.frame %>% select(-source,  -index) %>% rename(vignette=name), #%>% 
    #transform(n.vignettes=pp(n.vignettes), n.lines=pp(n.lines)), 
  row.names=FALSE)
```
