---
title: "A Look at CRAN Vignettes"
author: "Konrad Siek <siekkonr@fit.cvut.cz>"
date: "22 August 2017"
output: html_document
---

<style>
  .col2 {
    columns: 2 200px;         /* number of columns and width in pixels*/
    -webkit-columns: 2 200px; /* chrome, safari */
    -moz-columns: 2 200px;    /* firefox */
  }
  
   div.together {
    display: inline-block;
    width: 100%;
 }
</style>

```{r setup, include=FALSE}
library(ggplot2)
library(tibble)
library(dplyr)

pp <- function(number) format(number, big.mark=",", scientific=FALSE, trim=FALSE, digits=2)

#summary.vignettes <- readRDS("summary_vignettes.rds")
#summary.packages <- readRDS("summary_packages.rds")

summary.vignettes <- readRDS("~/workspace/R-dyntrace/summary_vignettes.rds")
summary.packages <- readRDS("~/workspace/R-dyntrace/summary_packages.rds")

top.100.packages <- scan("~/workspace/R-dyntrace/top_100_CRAN_packages.txt", as.character())
```

## Basic information

Number of packages: `r length(summary.packages$package)`  
Number of vignettes: `r length((summary.vignettes %>% filter(!is.na(index)))$name)`  
Lines of code in vignettes: `r pp(sum(summary.packages$n.lines)/1000)`KLOC

## Statistics

<div class="col2">
<div class="together">
```{r has-vignettes, echo=FALSE, comment="", fig.width=6, fig.height=6}
histogram <- summary.packages %>% mutate(has.vignettes=n.vignettes>0) %>% group_by(has.vignettes) %>% summarise(n.packages=n()) %>% mutate(percent=100*n.packages/sum(n.packages)) %>% arrange(n.packages)

labels <- histogram$has.vignettes
values <- histogram$n.packages
#par(mar=c(2,2,2,2))

pie(values, labels, main="Package has vignettes")
print(histogram %>% as.data.frame %>% transform(n.packages=pp(n.packages), percent=pp(percent)), row.names=FALSE)
```
</div>

```{r has-runnable-code, echo=FALSE, comment="", fig.width=6, fig.height=6}
histogram <- 
  summary.packages %>% 
  mutate(runnable.code=ifelse(n.vignettes == 0, "no vignettes",
                              ifelse(n.lines == 0, "no code",
                                     ifelse(n.lines <= 10, "≤ 10 lines",
                                            ifelse(n.lines <= 100, "≤ 100 lines", "> 100 lines"))))) %>%
  group_by(runnable.code) %>% summarise(n.packages=n()) %>% mutate(percent=100*n.packages/sum(n.packages)) %>% 
  arrange(n.packages)

labels <- histogram$runnable.code
values <- histogram$n.packages
#par(mar=c(2,2,2,2))

pie(values, labels, main="Package has runnable code in vignettes")
print(histogram %>% as.data.frame %>% transform(n.packages=pp(n.packages), percent=pp(percent)), row.names=FALSE)
```

</div>

```{r n-vignettes, echo=FALSE, comment="", fig.width=12, fig.height=5}
histogram <- summary.packages %>% group_by(n.vignettes) %>% summarise(n.packages=n()) %>% mutate(percent=100*n.packages/sum(n.packages))

labels <- histogram$n.vignettes
values <- histogram$n.packages
#par(mar=c(2,2,2,2))

barplot(values, names.arg=labels, space=0, main="Number of vignettes per package")
print(histogram %>% as.data.frame %>% transform(n.packages=pp(n.packages), percent=pp(percent)), row.names=FALSE)
```

## Top 100 packages

```{r top-100, echo=FALSE, comment="", fig.width=12, fig.height=5}
summary.top.100.packages <- tibble(package=top.100.packages) %>% left_join(summary.packages, by="package")
summary.top.100.vignettes <- tibble(package=top.100.packages) %>% left_join(summary.vignettes, by="package")
```

### Basic information

Number of packages: `r length(summary.top.100.packages$package)`  
Number of these actually installed: `r length((summary.top.100.packages %>% filter(!is.na(n.vignettes)))$package)`  
Number of vignettes: `r length((summary.top.100.vignettes %>% filter(!is.na(index)))$name)`  
Lines of code in vignettes: `r pp(sum((summary.top.100.packages %>% filter(!is.na(n.lines)))$n.lines)/1000)`KLOC

### Statistics

<div class="col2">
<div class="together">
```{r top-100-has-vignettes, echo=FALSE, comment="", fig.width=6, fig.height=6}
histogram <- summary.top.100.packages %>% mutate(has.vignettes=n.vignettes>0) %>% group_by(has.vignettes) %>% summarise(n.packages=n()) %>% mutate(percent=100*n.packages/sum(n.packages)) %>% arrange(n.packages)

labels <- as.logical(histogram$has.vignettes)
values <- histogram$n.packages
#par(mar=c(2,2,2,2))

pie(values, labels, main="Package has vignettes")
print(histogram %>% as.data.frame %>% transform(n.packages=pp(n.packages), percent=pp(percent)), row.names=FALSE)
```
</div>

```{r top-100-has-runnable-code, echo=FALSE, comment="", fig.width=6, fig.height=6}
histogram <- 
  summary.top.100.packages %>% 
  mutate(runnable.code=ifelse(n.vignettes == 0, "no vignettes",
                              ifelse(n.lines == 0, "no code",
                                     ifelse(n.lines <= 10, "≤ 10 lines",
                                            ifelse(n.lines <= 100, "≤ 100 lines", "> 100 lines"))))) %>%
  group_by(runnable.code) %>% summarise(n.packages=n()) %>% mutate(percent=100*n.packages/sum(n.packages)) %>% 
  arrange(n.packages)

labels <- histogram$runnable.code
values <- histogram$n.packages
#par(mar=c(2,2,2,2))

pie(values, labels, main="Package has runnable code in vignettes")
print(histogram %>% as.data.frame %>% transform(n.packages=pp(n.packages), percent=pp(percent)), row.names=FALSE)
```

</div>

```{r top-100-n-vignettes, echo=FALSE, comment="", fig.width=12, fig.height=5}
histogram <- summary.top.100.packages %>% group_by(n.vignettes) %>% summarise(n.packages=n()) %>% mutate(percent=100*n.packages/sum(n.packages))

labels <- histogram$n.vignettes
values <- histogram$n.packages
#par(mar=c(2,2,2,2))

barplot(values, names.arg=labels, space=0, main="Number of vignettes per package")
print(histogram %>% as.data.frame %>% transform(n.packages=pp(n.packages), percent=pp(percent)), row.names=FALSE)
```

### Data (top 100)

```{r top-100-data-packages, echo=FALSE, comment=""}
print(
   summary.top.100.packages %>% 
    as.data.frame %>% 
    transform(n.vignettes=pp(n.vignettes), n.lines=pp(n.lines)), 
  row.names=FALSE)
```

```{r top-100-data-vignettes, echo=FALSE, comment=""}
print(
   summary.top.100.vignettes %>% 
    as.data.frame %>% select(-source,  -index) %>% rename(vignette=name), #%>% 
    #transform(n.vignettes=pp(n.vignettes), n.lines=pp(n.lines)), 
  row.names=FALSE)
```

## Source code

```{r code, eval=FALSE}
library(dplyr)

packages <- installed.packages()[, "Package"]

filter_irrelecant_code <- function(source) {
  content <- source 
  
  index.el <- content %>% grepl("^[ \t]*$", .) # Remove empty lines
  content <- content[!index.el]
  
  index.cm <- content %>% grepl("^#", .) # Remove comments
  content <- content[!index.cm]
  
  content
}

count_source_lines <- function(paths) {
  result <- 0
  for (path in paths)
    result <- result + length(filter_irrelecant_code(path))
}

package_vignette_summary <- function(vignette_summary_data=NA) {
  data <- if(suppressWarnings(is.na(vignette_summary_data))) vignette_summary() else vignette_summary_data 
  data %>% 
    group_by(package) %>% 
    summarise(n.vignettes=sum(as.integer(!is.na(name))), n.lines=sum(ifelse(is.na(lines), 0, lines)))
}

vignette_summary <- function() {
  result.vignettes <- tibble(index=list(), package=list(), name=list(), source=list(), lines=list())
  first = TRUE
  
  for(package in packages) {
    vignettes <- vignette(package = package)$results
    
    if (length(vignettes) == 0)
      next
    
    vignettes.information <- 
      tibble(
          index=1:length(vignettes[, "Item"]), 
          package = vignettes[, "Package"], 
          name = vignettes[, "Item"]) %>%
      rowwise %>% 
      mutate(
        source={
          vignette.info <- vignette(name, package = package);  
          file.path(vignette.info$Dir, "doc", vignette.info$R)}) %>%
        mutate(lines = length(filter_irrelecant_code(readLines(source))))
    
    if (first) {
      result.vignettes <- vignettes.information
      first <- FALSE
    } else 
      result.vignettes <- rbind(result.vignettes, vignettes.information)
  }
  
  tibble(package=packages) %>% left_join(result.vignettes, by="package")
}

# On server
# library(tibble)
# summary.vignettes <- vignette_summary()
# summary.packages <- package_vignette_summary(summary.vignettes)
# saveRDS(summary.vignettes, "summary_vignettes.rds")
# saveRDS(summary.packages, "summary_vignettes.rds")

# On lappy
# library(tibble)
# summary.vignettes <- readRDS("summary_vignettes.rds"))
# summary.packages <- readRDS("summary_packages.rds"))
```



